---
title: "Which Factors Influence Canadians' Life Satisfaction?"
author: "Siyuan Tao"
date: 2015-07-23T21:13:14-05:00
categories: ["R"]
tags: ["R Markdown", "plot", "regression"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

## Title of your Report

# Siyuan Tao
# 2020-10-19

## Abstract

Here is where you give a brief (one paragraph overview of your entire paper). This should include some background/introduction, some methodology, results and conclusions.

## Introduction
Life satisfaction is an extensive term associated with everyone in the world. It is the evaluation of people's life (Ackerman, 2020). More specifically, it measures whether people feel satisfied with the lives they experienced. There are a variety of studies that tried to reveal the factors that influenced life satisfaction. For example, Baird, Lucas, and Donnellan (2010) found that age was associated with life satisfaction; life satisfaction decreased when people passed their 70s. Another study showed that people's life satisfaction increased as the household income increased (Frijters, Haisken-DeNew & Shields, 2004). 

However, even though many studies have focused on life satisfaction, few studies have focused on Canada and tested Canadians' life satisfaction. To fill the gap, this report will analyze factors associated with Canadians' life satisfaction using the data from the 2017 Canadian General Social Survey (GSS). Based on the comprehensive information from GSS, this report aims to test whether the factors influencing life satisfaction from previous studies could be applied to Canada and explore some other factors linked with Canadians' life satisfaction.

## Data
The data used in the report comes from the 2017 Canadian General Social Survey (GSS). GSS is a fundamental social survey in Canada. It collects Canadians' information on a diverse range of scale, including age, sex, income, and health condition (Statistics Canada, 2017). More importantly, the 2017 GSS includes the variable "feelings life," which is perfectly suitable for the report's research questions. Because it consists of this comprehensive information about Canadians that allows analyzing the relationship between life satisfaction and other factors, we chose GSS as the primary data to study in the report. There is a summary of the 2017 GSS below.

The target population of the 2017 Canadian GSS is "all non-institutionalized persons 15 years of age or older, living in the 10 provinces of Canada (Statistics Canada, 2019)." The total Canadian population in 2017 was 36,545,295, the population of ages between 0 to 14 years was 5,912,118 (Statistics Canada, 2020), so the population of people 15 years of age or older was 30,633,177. Thus, we can assume that the target population of the 2017 Canadian GSS is 30,633,177. Besides, the 2017 GSS contains 20,602 observations, which means the sample size of the data is 20,602.

Even though the GSS is suitable for the research questions, it has some drawbacks. First, the GSS collected data using Computer Assisted Telephone Interviewing (CATI), in which staff call the selected participants and collect information via phones (Statistics Canada, 2017). This way can help save money and energy, but it has some limitations. For example, researchers could only approach people who have phones. These people are usually associated with higher socioeconomic status and better education. Thus, not everyone in the population has the same probability of getting into the study, which may potentially make the result biased. Second, the GSS is usually collected over 6 to 12 months (Statistics Canada, 2017). People may experience different things during this long period, making their answers to the survey limited to timeliness. For example, people may give different answers about their mental health in a different time.  

## Model

To explore factors influncing Canadians' life satisfaction, we built a multiple linear regression model. Multiple linear regression model can help to identify the influences of multiple variables on a specific prediction variable. For example, if we want to use a multiple linear regression model to estimate this: $$y_{i} = \beta_{0} + \beta_{1}x_{1i} + \beta_{2}x_{2i} + \cdots + \beta_{k}x_{ki} + \epsilon_{i} $$

The multiple linear regression model should be: $$\hat{y} = b_{0} + b{1}x_{1} + b{2}x_{2} + \cdots + b_{k}x_{k}$$

Here, we use $b_{0} + b{1}x_{1} + b{2}x_{2} + \cdots + b_{k}x_{k}$ to predict $y_{i}$, so we use $\hat{y}$ to represent the prediction of $y_{i}$. Besides, $b_{0}$ is the intercept of the model, while $b_{1}, b_{2}, \dots,b_{k}$ refer to slopes corresponding to certain variables. 

In this case, we used variables including age, gender, family incomes, number of children, number of marriages, self rated health, self rated mental health, and whether living with a parter, to predict people's life satisfaction. Our model looks like this:$$\hat{Life Satisfaction} = b_{0} + b_{1}{Age} + b_{2}{Gender} + b_{3}{Family Income} + $$
$$b_{4}{Children Numbers} + b_{5}{Marriage Numbers} + b_{6}{Self Rated Health} +$$
$$b_{7}{Self Rated Mental Health} + b_{8}{Living with a Partner} $$

Here, the intercept $b_{0}$ means that if all variables are 0, then life satisfaction is expected to be $b_{0}$. The slope, for example, $b_{1}$, means that if all other variables are 0, when age increases for one unit, life satisfaction is expected to increase $b_{1}$.

In addition, some variables in the model are categorical variables, which means observations of these variables fall into different categories, and these categories are not continuous like numbers. For example, there are 6 levels of Family Income, which are "Less than \$25,000," "\$25,000 to \$49,999," "\$50,000 to \$74,999," "\$75,000 to \$99,999 ," "\$100,000 to \$124,999," and "\$125,000 and more." Similarly, variables of Health and Mental Health are also categorical variables. These two variables both have 6 levels, which are "Don't Know," "Poor," "Fair," "Good," "Very Good," and "Excellent." Participants will choose the category that best describes them. Some other categorical variables in the model include Gender, and Live with Partner. These two variables only include 2 different categories, which are "Male" or "Female," and "Yes" or "No," correspondingly. 

In this report, we used the language **R** to build the multiple regression model. Also, we regularized the model using the finite population correction factor. The data were collected without replacement, which means that when one person filled the survey, they would no longer be in the pool of selection. In this case, the standard error of each variable will be large and make the result of the model not valid (Glen, 2016). Thus, we included the finite population correction factor in the model, preventing the model from suffering the effect of sampling without replacement. 

## Results

The summary of the multiple linear regression is presented below:
```{r , echo=FALSE, message=FALSE, warning=FALSE}
# Import required packages
library(survey)
library(tidyverse)
```
```{r , echo=FALSE}
# Using population size and sample size to get the population correction factor
n = 20602
N=  30633177
fpc.srs = rep(N, n)
# Read the 2017 Canadian GSS dataset
gss <- read.csv("gss.csv")
GSS.design <- svydesign(id=~1, data=gss, fpc = fpc.srs)
# Build the multiple linear regression model
GSSanalysis <- svyglm(feelings_life ~ is_male + age + income_family + total_children + 
                        number_marriages + self_rated_health + self_rated_mental_health + 
                        lives_with_partner, GSS.design)
# Get the summary of the model
summary(GSSanalysis)
```

Based on the summary, we can build the multiple linear regression model to predict Canadians' life satisfaction: $$\hat{Life Satisfaction} = 7.200 + 0.006{Age} - 0.102{Male} + 0.018{<$125,000 and more>}  $$
$$-0.263{<$25,000 to $49,999>} - 0.142{<$50,000 to $74,999>}  $$
$$-0.106{<$75,000 to $99,999>} -0.400{<Less than $25,000>} + 0.067{ChildrenNumbers} $$
$$+  0.072{MarriageNumbers} +0.852{ExcellentHealth} +  0.655{VeryGood Health} $$
$$ +0.443{Good Health} +0.080{FairHealt}  -0.760{PoorHealth} $$
$$+ 0.679{ExcellentMental} + 0.224{VeryGoodMental} - 0.261{GoodMental}  $$
$$ - 1.289{FairMental}-2.575{PoorMental} + 0.091{Live With Partner} $$

By using the model, we can predict a person's life satisfaction based on their information. For example, if a 40-year-old male has 2 children, 1 marriage, lives with his partner, and rates himself as "good health" and "poor mental health," and the family income is in between \$75,000 and \$99,999, then the person's expected life satisfaction based on the model is:  $$\hat{Life Satisfaction} = 7.200 + 0.006 * 40 - 0.102 * 1 -0.106 * 1  + 0.067 * 2 +  0.072 * 1 $$
$$+0.443 * 1 -2.575 * 1 + 0.091 * 1 $$
$$ = 4.1582

Here you will include all results. This includes descriptive statistics, graphs, figures, tables, and model results. Please ensure that everything is well formatted and in a report style. You must also provide an explanation of the results in this section. You can overflow to an Appendix if needed. 

Please ensure that everything is well labelled. So if you have multiple histograms and plots, calling them Figure 1, 2, 3, etc. and referencing them as Figure 1, Figure 2, etc. in your report will be expected. The reader should not get lost in a sea of information. Make sure to have the results be clean, well formatted and digestible.

## Discussion

Here you will discuss conclusions drawn from the results and comment on how it relates to the original goal of the study (which was specified in the Introduction).

# Weaknesses

Here we discuss weaknesses of the study, data, analysis, etc. You can also discuss areas for improvement.

# Next Steps

Here you discuss subsequent work to be done after this report. This can include next steps in terms of statistical analysis (perhaps there is a more efficient algorithm available, or perhaps there is a caveat in the data that would allow for some new technique). Future steps should also be specified in terms of the study setting (eg. including a follow-up survey on something, or a subsequent study that would complement the conclusions of your report).


## References

